### 2015-06-13 writing/S'15 LSTM/abstract.tex
Paper trim and proof, new figures.

\begin{abstract}
Tree-structured neural networks aim to deliver a robust and principled method for representing sentence meaning, but these models largely have not outperformed simpler sequence-based models by substantial margins. We hypothesize that sequence models like LSTMs are able to discover and implicitly use the same kinds of recursive compositional structures that the tree-structured are built around, mitigating the advantage of the tree-structured models. We investigate this possibility by evaluating both models on an artificial task for which recursive compositional structure is crucial, and find that the sequence model is able to identify and exploit the underlying structure, though that it is less efficient at learning than the tree-structured models, only succeeding after exposure to a larger and richer set of training data.
\end{abstract}

>>>>

\begin{abstract}
Tree-structured neural networks aim to deliver a robust and principled method for representing sentence meaning, but these models largely have not outperformed simpler sequence-based models by substantial margins. We hypothesize that sequence models like LSTMs are able to discover and implicitly use the same kinds of recursive compositional structures that the tree-structured are built around, mitigating the advantage of the tree-structured models. We investigate this possibility by evaluating both models on an artificial task for which recursive compositional structure is crucial, and find that the sequence model is able to identify and exploit the underlying structure, though it is less efficient at learning than the tree-structured models, only succeeding after exposure to a larger and richer set of training data.
\end{abstract}

### 2015-06-13 writing/S'15 LSTM/data.tex
Paper trim and proof, new figures.

The data we use are labeled with the seven relation system of \newcite{maccartney2009extended}, depicted in Table \ref{b-table}.

\begin{table}[tp]
  \centering\small
  \renewcommand{\arraystretch}{1}
  \begin{tabular}{l c l l} 
    \toprule
    Name & Symb. & Set-theoretic definition \\ 
    \midrule
strict entailment         & $\natfor$   & $x \subset y$  \\ 
    strict rev. entailment & $\natrev$   & $x \supset y$  \\ 
    equivalence        & $\nateq$    & $x = y$   \\ 
    alternation        & $\natalt$   & $x \cap y = \emptyset \wedge x \cup y \neq \mathcal{D}$ \\ 
    negation           & $\natneg$   & $x \cap y = \emptyset \wedge x \cup y = \mathcal{D}$   \\
    cover              & $\natcov$   & $x \cap y \neq \emptyset \wedge x \cup y = \mathcal{D}$ \\ 
    independence       & $\natind$   & (else)\\
    \bottomrule
  \end{tabular}
  \caption{\label{b-table}MacCartney's seven mutually exclusive relations are defined abstractly on pairs of sets drawing from the universe $\mathcal{D}$, but can be straightforwardly applied to any pair of natural language words, phrases, or sentences.
  } %-%
\end{table}

@@

The data comes with parentheses representing complete binary bracketing. Our two models use this information differently. For the tree structured model, we assumble the tree structure for each sentence according to the parentheses, but do not treat them as word tokens. For the sequence model, we treat the parentheses as words, and learn their embeddings. This provides the two models with equivalent data, so that their ability to handle unseen structures can be fairly compared.

\paragraph{The corpus}
The sentence pairs in the corpus are divided into thirteen bins according to the number of logical connectives (\word{and, or, not}) in the longer of the two sentences in the pair. We test the model on each bin separately (58k total examples, using an 80/20\% train/test split) in order to evaluate how its performance depends on the complexity of the sentences. In three experiments, we train our models on the training portions of bins 0--3 (62k examples), bins 0--4 (90k), and bins 0--6 (160k). Capping the size of the training sentences below length 12 allows us to evaluate how the models interpret the sentences: if their performance falls off abruptly above the cutoff, it is reasonable to assume that the models are depending heavily on specific sentence structures, and cannot generalize to new structures. If then performance decays gradually\footnote{Since sentences are fixed-dimensional vectors of fixed-precision floating point numbers, the models will inevitably will start to make errors on sentences above some length, and the (necessary) use of L2 regularization exacerbates this for sentences which are longer than those seen in training.} with no such abrupt change, then it must have learned a more generally valid interpretation function for the language which respects its recursive structure.

>>>>

The data we use are labeled with the seven mutually exclusive logical reations of \newcite{maccartney2009extended}, which distinguish entailment in two directions ($\natfor$, $\natrev$), equivalence ($\nateq$), two types of contradiction ($\natneg$, $\natalt$), and two types of semantic independence ($\natind$, $\natcov$).

%\begin{table}[tp]
%  \centering\small
%  \renewcommand{\arraystretch}{1}
%  \begin{tabular}{l c l l} 
%    \toprule
%    Name & Symb. & Set-theoretic definition \\ 
%    \midrule
%strict entailment         & $\natfor$   & $x \subset y$  \\ 
%    strict rev. entailment & $\natrev$   & $x \supset y$  \\ 
%    equivalence        & $\nateq$    & $x = y$   \\ 
%    alternation        & $\natalt$   & $x \cap y = \emptyset \wedge x \cup y \neq \mathcal{D}$ \\ 
%    negation           & $\natneg$   & $x \cap y = \emptyset \wedge x \cup y = \mathcal{D}$   \\
%    cover              & $\natcov$   & $x \cap y \neq \emptyset \wedge x \cup y = \mathcal{D}$ \\ 
%    independence       & $\natind$   & (else)\\
%    \bottomrule
%  \end{tabular}
%  \caption{\label{b-table}MacCartney's seven mutually exclusive relations are defined abstractly on pairs of sets drawing from the universe $\mathcal{D}$, but can be straightforwardly applied to any pair of natural language words, phrases, or sentences.
%  } %-%
%\end{table}

@@

The data come with parentheses representing complete binary bracketing. Our two models use this information differently. For the tree structured model, we assumble the tree structure for each sentence according to the parentheses, but do not treat them as word tokens. For the sequence model, we treat the parentheses as words, and learn their embeddings. This provides the two models with equivalent data, so that their ability to handle unseen structures can be fairly compared.

\paragraph{The corpus}
The sentence pairs in the corpus are divided into thirteen bins according to the number of logical connectives (\word{and, or, not}) in the longer of the two sentences in the pair. We test the model on each bin separately (58k total examples, using an 80/20\% train/test split) in order to evaluate how its performance depends on the complexity of the sentences. In three experiments, we train our models on the training portions of bins 0--3 (62k examples), 0--4 (90k), and 0--6 (160k). Capping the size of the training sentences below length 12 allows us to evaluate how the models interpret the sentences: if their performance falls off abruptly above the cutoff, it is reasonable to assume that the models are depending heavily on specific sentence structures, and cannot generalize to new structures. If then performance decays gradually\footnote{Since sentences are fixed-dimensional vectors of fixed-precision floating point numbers, the models will inevitably will start to make errors on sentences above some length, and the (necessary) use of L2 regularization exacerbates this for sentences which are longer than those seen in training.} with no such abrupt change, then it must have learned a more generally valid interpretation function for the language which respects its recursive structure.

### 2015-06-13 writing/S'15 LSTM/discussion.tex
Paper trim and proof, new figures.

The results are shown in Figure \ref{prop-results}. All three models were able to largely memorize the training data (TODO, TODO, and TODO for $\le$6). With sufficient training data (in the $\le$6 setting), the two tree models were able to perform at above 99\% (TODO) on examples of size $\le$2 with a steady decay in performance continuing through TODO\% at size 12, and similar, if steeper, decays with smaller training sets. These robustly beat the simple baselines reported in Bowman et al.: the most frequent class occurs just over 50\% of the time, and a summing model that ignores word order does reasonably on the shortest examples but falls below 60\% by bin 4.

The LSTM performs fairly poorly in the $\le3$ setting, with performance at 3 a mediorce TODO\%, and an abrubt drop to TODO\% at 4, suggesting that the model did not acquire the needed generalizations. With more ample training data in the $\le6$ condition, though, the LSTM shows fairly good accuracy on short examples, and a smooth decay with no abrupt drop.

@@

We find that all three \todo{Still using three?} models can exploit recursive structure to interpret sentences with complex unseen structures, and that tree structured model's biases allow them to learn to do this more effectively from less data. 
We interpret these results as evidence that the similar performance of tree and sequence models can be attributed to their fundamentally similar representational systems, with both models recognizing compositional structure when it is present, the former doing so more reliably, and the latter benefiting elsewhere---either because of architectural choices that make backpropagation more effective, or because interpreting sentences from left to right allows the model to better simulate other aspects of human language comprehension. We finally suggest that there is value in attempting to further develop this ability in sequence models without fully sacrificing the flexibility that makes these models succeed.

>>>>

The results are shown in Figure \ref{prop-results}\todo{Next version of figures will have vertical dividing lines at 3,4,6}. All four models were able to largely memorize the training data (TODO, TODO, TODO, and TODO for $\le$6). In the $\le$6 setting, the three tree models were able to perform at above TODO\% on examples of size $\le$2 with a steady decay in performance continuing through TODO\% at size 12, and similar, if steeper, decays with smaller training sets. All four robustly beat the simple baselines reported in Bowman et al.: the most frequent class occurs just over 50\% of the time, and a summing model that ignores word order does reasonably on the shortest examples but falls below 60\% by bin 4.

The LSTM performs fairly poorly in the $\le3$ setting, with performance at 3 a mediocre TODO\%, and an abrupt drop to TODO\% at 4, suggesting that the model did not acquire the needed generalizations. With more ample training data in the $\le6$ condition, though, the LSTM shows fairly good accuracy on short examples, and a smooth decay with no abrupt drop.

[Something about the learning curve that doesn't involve a figure. Something about the learning curve that doesn't involve a figure. Something about the learning curve that doesn't involve a figure.]

@@

We find that all four models can exploit recursive structure to interpret sentences with complex unseen structures, and that tree structured model's biases allow them to learn to do this more effectively from less data. 
We interpret these results as evidence that the similar performance of tree and sequence models can be attributed to their fundamentally similar representational systems, with both models recognizing compositional structure when it is present, the former doing so more reliably, and the latter benefiting elsewhere---either because of architectural choices that make backpropagation more effective, or because interpreting sentences from left to right allows the model to better simulate other aspects of human language comprehension. We finally suggest that, because of the well supported linguistic claim that the kind of recursive structure that we study here is key to language understanding, there is likely to be value in developing sequence models that can more efficiently exploit this structure without fully sacrificing the flexibility that makes these models succeed.

### 2015-06-13 writing/S'15 LSTM/methods.tex
Paper trim and proof, new figures.

We use the model architecture depicted in Figure \ref{fig:model:top}. This approach builds on the model used in \newcite{Bowman:Potts:Manning:2014}. The model architecture uses two copies of a single sentence embedding network---a TreeRNN, TreeRNTN, or LSTM RNN, depending on the precise experiment---to embed the premise and hypothesis sentences, and then uses those embeddings as the features for a multilayer classifier which predicts one of the seven relations. Since the embeddings are computed separately, the embedding models are forced to encode every element of the sentence meaning that the downstream model will need.

\paragraph{Classifier}
The classifier component of the model consists of a combinging layer which takes the two sentence representations as inputs, followed by two neural network layers, then a softmax classifier.
For the combining layer, we use a (recursive) neural tensor network ((R)NTN, \citealt{chen2013learning}) layer, which sums the output of a plain recursive/recurrent neural network layer with a vector computed using two multiplications with a learned (full rank) third-order tensor parameter:
\begin{gather} 
\label{TreeRNN}
\vec{y}_{\textit{RNN}} = \tanh(\mathbf{M} \colvec{2}{\vec{x}^{(l)}}{\vec{x}^{(r)}} + \vec{b}\,) \\
\label{TreeRNTN} 
\vec{y}_{\textit{RNTN}} = \vec{y}_{\textit{RNN}} + \tanh(\vec{x}^{(l)T} \mathbf{T}^{[1 \ldots n]} \vec{x}^{(r)})
\end{gather}

@@

\paragraph{Sentence embedding}
The sentence embedding component of the model transforms the (jointly learned) embeddings of the input words into a single sentence vector. We experiment with two tree-structured models (Figure \ref{fig:model:tree}), one with an RNTN activation function or a plain RNN activation function. Both are structured using the parses that are provided with the data. In addition, we use a sequence model (Figure \ref{fig:model:seq}) with an LSTM activation function \cite{hochreiter1997long} implemented as in \newcite{zaremba2015recurrent}.\todo{Okay to omit negative result on sequence RNN?}

>>>>

We use the model architecture depicted in Figure \ref{fig:model:top}, which builds on the model used in \newcite{Bowman:Potts:Manning:2014}. The model architecture uses two copies of a single sentence embedding network (a tree or sequence model) to embed the premise and hypothesis (left and right side) sentences, and then uses those embeddings as the features for a multilayer classifier which predicts one of the seven relations. Since the embeddings are computed separately, the embedding models are forced to encode every element of the sentence meaning that the downstream model will need.

\paragraph{Classifier}
The classifier component of the model consists of a combinging layer which takes the two sentence representations as inputs, followed by two neural network layers, then a softmax classifier.
For the combining layer, we use a neural tensor network (NTN, \citealt{chen2013learning}) layer, which sums the output of a plain recursive/recurrent neural network layer with a vector computed using two multiplications with a learned (full rank) third-order tensor parameter:
\begin{gather} 
\label{TreeRNN}
\vec{y}_{\textit{NN}} = \tanh(\mathbf{M} \colvec{2}{\vec{x}^{(l)}}{\vec{x}^{(r)}} + \vec{b}\,) \\
\label{TreeRNTN} 
\vec{y}_{\textit{NTN}} = \vec{y}_{\textit{RNN}} + \tanh(\vec{x}^{(l)T} \mathbf{T}^{[1 \ldots n]} \vec{x}^{(r)})
\end{gather}

@@

\paragraph{Sentence embedding}
The sentence embedding component of the model transforms the (jointly learned) embeddings of the input words into a single sentence vector. We experiment with tree-structured models (Figure \ref{fig:model:tree}) with TreeRNN (eqn.~\ref{TreeRNN}), TreeRNTN (eqn.~\ref{TreeRNTN}), and TreeLSTM \cite{tai2015improved} activation functions. In addition, we use a sequence model (Figure \ref{fig:model:seq}) with an LSTM activation function \cite{hochreiter1997long} implemented as in \newcite{zaremba2015recurrent}. Experiments with simpler non-LSTM RNN sequence models tended to underfit the training data, and are not included here.
